Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2025-04-29 Latex_files/Thesis/chapters/chapter 1/main.tex
updates 25/04/25

\chapter{Introduction}
Crystallisation is one of the oldest forms of phase 
separation used by humanity \cite{Schoen1956}, put 
simply, it is the formation and growth of a new 
structured phase within a disordered bulk phase. 
This has applications in a number of industries such 
as pharmaceuticals \cite{Gao2017}, food production 
\cite{Hartel2002}, and electronics \cite{Myerson2002}. 
Where the extraction of dilute materials can help 
improve product quality while keeping production costs 
low. The industrialisation of crystallisation has 
allowed engineers to reliably and efficiently induce 
the crystal formation within a bulk phase. However, 
while on a large scale crystallisation is seemingly an 
understood physical process, just a small amount of 
investigation into the literature reveals that at a 
micro scale there is yet to be unifying theory that can 
accurately explain the process of crystallisation 
\cite{Fu2021}. Crystallisation can be subdivided into 
two steps: Nucleation and Crystal growth. The latter 
focusing on how an already stable crystal grows and how 
it takes on its final shape. Whereas the former is more 
concerned with the first initial moments of a crystals 
formation within a bulk phase.

In this chapter we will outline our current understanding 
of nucleation, why there is a gap in the literature, and why 
there is still a need for local control of nucleation events. 
Furthermore, we will also highlight recent developments 
involving the use of optical tweezers (and other laser based 
methods) in order to develop our understanding of nucleation. 
Lastly we highlight the connection between optical tweezer's 
ability to probe fluid viscosity via rotational motion, and 
the known link between nucleation and fluid flow, to propose 
a novel potential method of creating localised nucleation events.

\section{Nucleation}
Nucleation is an example of a binary phase separation, 
where a dilute phase is miscible in a bulk phase, more 
often called the solute and solvent respectively. When 
a small amount of solute is added to the solvent the 
solute will diffuse throughout the solvent, reaching
an equilibrium concentration throughout the solution. 
Because of molecular interactions, the two can only remain 
in equilibrium while below a specific concentration ($C_{eq}$) 
- below which the chemical potential $\mu$ for a single phase 
is greater than the potential required to separate the two. 
Once $C_{eq}$ is exceeded there is a chemical potential 
difference driving the solution to separate the two phases. 
Since different combinations of solute and solvent will 
have different equilibrium concentrations, researchers often 
instead measure the ratio between the solute and solvent by 
using 'supersaturation' \cite{Mullin2001}:
\begin{align}
	\label{eq:supersaturation}
	S(T) = \frac{C_{sol}}{C_{eq}(T)}
\end{align}

@@

\subsection{Primary \& Secondary nucleation}
From an industrial perspective, the nucleation process can 
be broadly categorised into either primary of secondary
nucleation. The former describes the formation of an initial
nucleus within the bulk phase, absent of any external stimuli.
Primary nucleation is therefore considered stochastic as there 
is no means of predicting where a nucleus may form, or how long 
it will take. The only reliable indicator being that higher 
supersaturations will result in faster nucleation rates. At a 
small scale one can estimate the nucleation rate by making 
repeated measurements of sample solutions and seeing how many 
have nucleated after a given time, giving us a Poisson probability distribution.
\begin{align}
	P(t) = 1 - exp\left[-JV(t-t_g)\right] = \frac{M^*(t)}{M}
	\label{eq:poisson}
\end{align}

Where $J$ is the nucleation rate, $t_g$ is the 'growth time', 
$V$ is the volume of the individual samples, and $M^*(t)$ \& 
$M$ are the number of nucleated samples and the total number 
of samples used respectively. The nucleation rate is one of 
the most important factors used in evaluating a crystallisers 
performance, providing a rough estimation of the final product 
mass over a given time frame. While \eqref{eq:poisson} is useful 
for studying the effects of different parameters at a small scale, 
for industrial applications there are to many external factors 
for primary nucleation to be measured accurately.

Secondary nucleation is the result of a initial seed crystal 
inducing further nucleation within the bulk solution 
\cite{Botsaris1976}. The seed crystals is often specially 
prepared ahead of time and added to a supersaturate solution. 
Due to interactions with the surrounding fluid and the container 
walls it is possible for the seed crystals to act as a surface
for further nucleation \cite{Anwar2015}. Control of secondary 
nucleation events are crucial for ensuring reliable industrial 
crystalliser performance. Fig.~\ref{fig:secondary} depicts an 
attempt of classifying every possible mechanism that could lead 
to secondary nucleation.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.95\linewidth]{secondary_nucleation.jpg}
	\caption{Secondary Nucleation mechanisms, classified by
		Agrawal and Paterson \cite{Agrawal2015}}
	\label{fig:secondary}
\end{figure}

@@

Where $S$ is the supersaturation from eq.~\eqref{eq:supersaturation} 
$v$ is the approximate volume of an individual molecule, 
$k_B$ is the Boltzmann constant, and $\sigma_{inf}$ is the 
interfacial tension of the bulk solution. This assumes that 
the nucleus will have a spherical morphology so that the 
surface tension $\sigma_{inf}$ is a scalar value. Looking at 
\eqref{eq:CNT} suggests that there must be some critical size 
$r$ where the free energy gain from the nucleus exceeds the 
surface tension of the surrounding fluid. This is reflected in 
fig.~\ref{fig:free_energy} where we plot the free energy of the 
system against nucleus size. This reveals a critical size above 
which the gain in free energy exceeds the interfacial tension. 
Furthermore fig.~\ref{fig:free_energy} shows how increasing the 
supersaturation of the system reduces said barrier. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{Free_Energy_Diagram.png}
	\caption{Free energy diagram of a newly formed nucleus according 
		     to the Classical Nucleation Theory. The total free energy (blue)
		     is due to the competition between the volume free energy gain
		     (green) and the interfacial free energy cost (orange). Dotted
		     lines are for a higher supersaturation than the solid lines,
		     the interfacial energy cost is independent of supersaturation.}
	\label{fig:free_energy}
\end{figure}

@@

CNT is often regarded as a good description of the macro system, 
its obvious that for all crystallization systems there is an 
inherent energy barrier that dictates the nucleation rate. 
Where it falters is in its predictive ability, both in estimating 
nucleation rates \cite{Gharibeh2005, Vekilov2010}, and in the 
structure of newly formed nuclei \cite{Lee1999, Yau2001}.
Recent studies suggest classical nucleation is merely one of 
many possible pathways that can be taken to produce an ordered 
crystalline phase. Prompting the development of alternative 
theories to better describe the nucleation process.

>>>>

\chapter{Introduction}
Crystallisation is one of the oldest forms of phase 
separation used by humanity \cite{Schoen1956}, put 
simply, it is the formation and growth of a new 
structured phase within a disordered bulk phase. 
This has applications in a number of industries such 
as pharmaceuticals \cite{Gao2017}, food production 
\cite{Hartel2002}, and electronics \cite{Myerson2002}. 
Where the extraction of dilute materials can help 
improve product quality while keeping production costs 
low. Modern crystallisation methods involve manipulating 
the solubility of the target chemical (aka the solute) 
in order to encourage the formation of a separate phase
that can be easily extracted \cite{Gao2017}. This is 
simple enough to understand on a macro scale when the 
concentration can be assumed as a time-dependent 
variable; but at a micro level the crystallisation process
is not merely subject to solubility but also due to 
the inter-molecular forces that can inhibit or enhance 
the formation of a new phase. Theories attempting to describe
the phenomena but none have proven viable to explain the 
various pathways available \cite{Fu2021}.

Crystallisation can be subdivided into two stages: The first 
stage - nucleation - is the formation of a brand new phase 
within the bulk solution, while nucleation can occur readily 
in most solutions it is difficult to predict if a nucleus 
will remain stable within the solution or if it will be 
dissolve. The latter stage - crystal growth - is focused 
instead on how an already stable nucleus grows into its 
final shape depending on the conditions of the surrounding
fluid, understanding how and where deformities occur within
the crystal structure are of extreme importance to several 
industries \cite{Gao2017,Myerson2002}.

This chapter will outline our current understanding of nucleation, 
where there is a gap in the literature, and why there is still a 
need for local control of nucleation events. Furthermore, it will 
also highlight recent developments involving the use of optical 
tweezers (and other \textit{in situ } methods) in order to develop 
our understanding of nucleation. Lastly we highlight the connection 
between optical tweezer's ability to probe fluid viscosity via 
rotational motion, and the known link between nucleation and fluid 
flow, to propose a novel potential method of creating localised 
nucleation events.

\section{Nucleation}
Nucleation is an example of a binary phase separation, 
where a dilute phase is miscible in a bulk phase, more 
often called the solute and solvent respectively. When 
a small amount of solute is added to the solvent the 
solute will diffuse throughout the solvent, reaching
an equilibrium concentration throughout the solution. 
Because of molecular interactions, the two can only remain 
in equilibrium while below a specific concentration ($C_{eq}$) 
- below which the chemical potential $\mu$ for a single phase 
is greater than the potential required to separate the two. 
Once $C_{eq}$ is exceeded there is a chemical potential 
difference driving the solution to separate the two phases.
This does not mean that nucleation will occur immediately 
once $C_{eq}$ is exceeded, only that there is a chemical
potential supporting nucleation.

Since different combinations of solute and solvent will 
have different equilibrium concentrations, researchers often 
instead measure the ratio between the solute and solvent by 
using 'supersaturation' \cite{Mullin2001}:
\begin{align}
	\label{eq:supersaturation}
	S(T) = \frac{C_{sol}}{C_{eq}(T)}
\end{align}

@@

\subsection{Primary \& Secondary nucleation}
nucleation process be broadly categorised into either primary 
or secondary nucleation. The former describes the formation 
of an initial nucleus within the bulk phase, absent of any 
external surfaces. Primary nucleation is dependent only on 
the local concentrations of the solute which can fluctuate 
unpredictably, as such it is often modelled at a large scale 
as a stochastic process. The only reliable factor being 
that higher supersaturations will result in more nucleation 
events to occur sooner.

An important metric for crystallisation procedures is the 
nucleation rate of the solution, a measure of the total 
volume or mass of crystalline material produced for a given 
set of initial conditions. At an industrial scale this 
metric is a contribution of both primary and secondary 
nucleation events. At a small scale one can estimate the 
contribution of the primary nucleation pathway by making 
repeated measurements of sample solutions and seeing how 
many have nucleated after a given time, giving us a Poisson 
probability distribution.
\begin{align}
	P(t) = 1 - exp\left[-JV(t-t_g)\right] = \frac{M^*(t)}{M}
	\label{eq:poisson}
\end{align}

Where $J$ is the nucleation rate, $t_g$ is the 'growth time' 
which accounts for the initial delay due to the stochastic 
nature of primary nucleation, $V$ is the volume of the individual 
samples, and $M^*(t)$ \& $M$ are the number of nucleated samples 
and the total number of samples used respectively. While 
\eqref{eq:poisson} is useful for studying the effects of different 
parameters at a small scale, for industrial applications there are 
to many external factors for primary nucleation to be measured 
accurately. Overall while the primary nucleation rate is significant 
when studying solutions for small volumes, the contributions of 
secondary nucleation is far more significant when considering 
nucleation rate for an industrial crystalliser.

Secondary nucleation is the result of a initial seed crystal 
inducing further nucleation within the bulk solution 
\cite{Botsaris1976}. The seed crystals are often prepared ahead 
of time and added to a supersaturate solution. Due to interactions 
with the surrounding fluid and the container walls it is possible 
for the seed crystals to act as a surface for further nucleation \cite{Anwar2015}. However this is just one possible way in which 
a seed crystal can enhance the nucleation rate, depending on the 
conditions of the crystalliser there are multiple mechanism by 
which the seed crystal can contribute to the nucleation rate of 
the system. Control of secondary nucleation events are crucial 
for ensuring reliable industrial crystalliser performance. Fig.~\ref{fig:secondary} depicts an attempt of classifying every 
possible mechanism that could lead to secondary nucleation.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.95\linewidth]{secondary_nucleation.jpg}
	\caption{Secondary Nucleation mechanisms, classified by
		Agrawal and Paterson \cite{Agrawal2015}}
	\label{fig:secondary}
\end{figure}

@@

Where $S$ is the supersaturation from eq.~\eqref{eq:supersaturation} 
$v$ is the approximate volume of an individual molecule, 
$k_B$ is the Boltzmann constant, and $\sigma_{inf}$ is the 
interfacial tension of the bulk solution. This assumes that 
the nucleus will have a spherical morphology so that the 
surface tension $\sigma_{inf}$ is a scalar value. Looking at 
\eqref{eq:CNT} suggests that there must be some critical size 
$r$ where the free energy gain from the nucleus exceeds the 
surface tension of the surrounding fluid. This is reflected in 
fig.~\ref{fig:free_energy} where we plot the free energy of the 
system against nucleus size. This reveals a critical size above 
which the gain in free energy exceeds the interfacial tension. 
Furthermore fig.~\ref{fig:free_energy} shows how increasing the 
supersaturation of the system reduces said barrier. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{Free_Energy_Diagram.png}
	\caption{Free energy diagram of a newly formed nucleus according 
		     to the Classical Nucleation Theory. The total free energy (blue)
		     is due to the competition between the volume free energy gain
		     (green) and the interfacial free energy cost (orange). Dotted
		     lines are for a higher supersaturation than the solid lines,
		     the interfacial energy cost is independent of supersaturation.
		     Diagram was constructed in python using a naive assumption 
		     of $v=1\times 10^-12\ m^3$ and $\sigma_{inf}=0.0001\ J/m^2$}
	\label{fig:free_energy}
\end{figure}

@@

CNT is often regarded as a good description of the macro system, 
its obvious that for all crystallization systems there is an 
inherent energy barrier that must be overcome prior to nucleation. 
Where it falters is in its predictive ability, both in estimating 
nucleation rates \cite{Gharibeh2005, Vekilov2010}, and in the 
structure of newly formed nuclei \cite{Lee1999, Yau2001}.
Recent studies suggest classical nucleation is merely one of 
many possible pathways that can be taken to produce an ordered 
crystalline phase. In situations where the solution is heavily 
saturated with solute the driving force promoting nucleation is 
so large that a nucleus can form in the bulk solution in a single 
step.

This is however not accurate at lower supersaturations, closer
to the saturation point the solvent's surface tension is so 
significant that the addition of individual solute molecules
should not be enough to overcome the dissolution of the nucleus. 
And yet nucleation is still possible, occurring far more reliably
than predicted by the CNT. This suggests that the formation of 
a nucleus is due to larger aggregates of solute that are stable 
within the bulk solution. This has been confirmed experimentally 
and lead to the development of Two Step Nucleation.

### 2025-04-29 Latex_files/Thesis/chapters/chapter 3/main.tex
updates to chp 4 & 2
updates 25/04/25

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Position detection methods}
In order to accurately capture the dynamics of a trapped particle, a 
position detection system is required. There are 3 possible methods 
of position detection: video-analysis, lateral-effect position sensing, 
and photodiodes.

@@

A quadrant photo diode (QPD) is a frequently used position detection 
system for optical tweezers due to their high sampling rate, high 
degree of precision, and ease of set up. The QPD is constructed of 
four photo diodes assembled in a quadrant formation, when a particle 
is trapped the interference pattern produced is focused onto the QPD, 
with the maximum intensity mapping to the particle's centre of mass. 
By summing the voltages of the horizontal and vertical quadrants together 
the particle's centre of mass is tracked in the x-y plane. Axial 
displacement can be estimated by observing the change in the total 
voltage of the QPD. The outputted signal gives an indication of the 
particle's relative displacement from the beam focus, but in order to 
convert the signal to distance units the trap needs to be calibrated 
(assuming a linear response curve).
\begin{figure}[h!]
	\centering
	\includegraphics[height=\linewidth, angle=270]{QPD_Lateral_effect.pdf}
	\caption{Comparison between QPD and Lateral effect photodiodes.
	The four quadrants of a QPD (left) experience different photocurrents
	based on the total intensity of light incident on each section 
	(labelled A, B, C, D). 
	Whereas a Lateral effect sensor (right) uses the resistive properties
	of the photodiode surface to vary the create different photocurrents 
	passing through the anodes A, B, C, and D.}
\end{figure}

A lateral-effect sensor has a similar output but works using a the 
entire sensor as a single cell analogous to the focal plane of the 
trapping beam. The four corners of the sensor act as anodes connected 
to a base plate cathode, as the beam moves across the surface of the 
detector each anode will experience a different photocurrent depending 
on how close the centre of the interference pattern is to each anode. 
The advantage of a lateral effect detector is that the linear regime 
is much larger than a QPD making it much better for monitoring the 
position of a trapped particle. However, Lateral-effect sensors are 
often limited in their spacial resolution due to high signal-to-noise 
ratios, requiring a high intensity of light on the sensor in order to 
get a clean signal. As a result, most optical force measurements are 
conducted using a QPD as opposed to a lateral-effect sensor, as often 
the displacement is small enough that the signal-displacement curve 
can be considered linear.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fourier Optics and 4f correlators}
\label{sec:4f_correlation}
As shown in \ref{fig:setup}, after the Galvano mirror there are a pair
of focal lenses ($f_3$ and $f_4$) that do not seem to serve a clear
purpose. These are in fact crucial for the operation of the Galvano, the
pair of them can be more accurately called a 4f correlator.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Synthesis of Birefringent Micro spheres}
\label{sec:vaterite}
There are several options for particles that can be rotated 
using optical tweezers \cite{Parkin2009, Saito2022}. Over 
the course of the project two different micro spheres where 
investigated, Vaterite and liquid crystal droplets. Both can 
be readily synthesised in the lab and are will rotate at a 
variety of sizes (see \ref{sec:opt_torque}).

Vaterite is a polymorph of calcium carbonate that is rarely 
seen in nature due to its low stability \cite{KonopackaLyskawa2019}. 
All three polymorphs are inherently birefringent meaning 
that they can be rotated using circularly polarised light. 
However unlike its other polymorphs of calcite and aragonite, 
when synthesised vaterite will typically form small spherical 
particles making them ideal for optical trapping and rotation. 
Synthesis of Vaterite micro spheres requires fine control of 
the crystal growth process in order to maintain polymorphic 
stability. Though for the purposes of optical rotation the 
exact polymorph is not as important as its morphology as all 
3 polymorphs are inherently birefringent.

Vaterite samples where made by the first preparing equal amounts 
of $CaCl_2$ and $Na_2CO_3$ at a concentration of $0.33M$, at 
the same time a vial of $0.33M\ MgSO_4$ was prepared and set 
aside for later. Firsta small vial was filled with $1.5mL$ of 
$CaCl_2$ followed by $60\mu L$ and $90\mu L$ of $MgSO_4$ and 
$NaCO_3$ respectively, forming a seed solution. Next, a larger 
vial was filled with $5\ mL$, $1.5\ mL$, and $1\ mL$ of $CaCL_2$, 
$MgSO_4$, and $NaCO_3$ respectively followed by the seed solution. 
After 10 minutes of slow but continuous mixing a few drops of 
Agepon was added to halt the reaction, the solution was filtered 
and washed 3 times with distilled water before being suspended 
in water.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Rotation of birefringent micro spheres} 
Optical tweezing has often been used for micro-rheology, by computing the
exact forces being exerted on the trapped sphere, one can determine the
local temperature/viscosity of the medium \cite{Millen2014, RodriguezSevilla2018}.
When it comes to optical rotation, a rotating particle will experience a
fluid drag torque that is proportional to its rotation rate, the maximum 
rotation rate is when the fluid drag is equal to the optical torque 
of the laser \cite{RodriguezSevilla2018}. Optical rotors have been used
to measure local temperature changes \cite{Millen2014}, or for understanding 
how fluid shear propagates through a medium \cite{Knoener2005}. Likewise, 
one can use a beam steering arrangement to probe the drag force of the 
fluid, by understanding the trap strength (calibrating using a low 
frequency signal) one can measure the drag force experienced by the local 
fluid \cite{RobertsonAnderson2018}. I

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Position detection}
\label{sec:position_detection}
In order to accurately capture the dynamics of a trapped particle, a 
position detection system is required. There are 3 possible methods 
of position detection: video-analysis, lateral-effect position sensing, 
and photodiodes.

@@

A lateral-effect sensor has a similar output but works using a the 
entire sensor as a single cell analogous to the focal plane of the 
trapping beam. The four corners of the sensor act as anodes connected 
to a base plate cathode, as the beam moves across the surface of the 
detector each anode will experience a different photocurrent depending 
on how close the centre of the interference pattern is to each anode. 
The advantage of a lateral effect detector is that the linear regime 
is much larger than a QPD making it much better for monitoring the 
position of a trapped particle. However, Lateral-effect sensors are 
often limited in their spacial resolution due to high signal-to-noise 
ratios, requiring a high intensity of light on the sensor in order to 
get a clean signal. As a result, most optical force measurements are 
conducted using a QPD as opposed to a lateral-effect sensor, as often 
the displacement is small enough that the signal-displacement curve 
can be considered linear.
\begin{figure}[h!]
	\centering
	\includegraphics[height=\linewidth, angle=270]{QPD_Lateral_effect.pdf}
	\caption{Comparison between QPD and Lateral effect photodiodes.
	The four quadrants of a QPD (left) experience different photocurrents
	based on the total intensity of light incident on each section 
	(labelled A, B, C, D). 
	Whereas a Lateral effect sensor (right) uses the resistive properties
	of the photodiode surface to vary the create different photocurrents 
	passing through the anodes A, B, C, and D.}
\end{figure}

A quadrant photo diode (QPD) is a frequently used position detection 
system for optical tweezers due to their high sampling rate, high 
degree of precision, and ease of set up. The QPD is constructed of 
four photo diodes assembled in a quadrant formation, when a particle 
is trapped the interference pattern produced is focused onto the QPD, 
with the maximum intensity mapping to the particle's centre of mass. 
By summing the voltages of the horizontal and vertical quadrants together 
the particle's centre of mass is tracked in the x-y plane. Axial 
displacement can be estimated by observing the change in the total 
voltage of the QPD. The outputted signal gives an indication of the 
particle's relative displacement from the beam focus, but in order to 
convert the signal to distance units the trap needs to be calibrated 
(assuming a linear response curve).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fourier Optics and 4f correlators}
\label{sec:4f_correlation}
As shown in \ref{fig:setup}, after the Galvano mirror there are a pair
of focal lenses ($f_3$ and $f_4$) that do not seem to serve a clear
purpose. These are in fact crucial for the operation of the Galvano, the
pair of them can be more accurately called a 4f correlator.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Synthesis of Birefringent Micro spheres}
\label{sec:vaterite}
There are several options for particles that can be rotated 
using optical tweezers \cite{Parkin2009, Saito2022}. Over 
the course of the project two different micro spheres where 
investigated, Vaterite and liquid crystal droplets. Both can 
be readily synthesised in the lab and are will rotate at a 
variety of sizes (see \ref{sec:opt_torque}).

Vaterite is a polymorph of calcium carbonate that is rarely 
seen in nature due to its low stability \cite{KonopackaLyskawa2019}. 
All three polymorphs are inherently birefringent meaning 
that they can be rotated using circularly polarised light. 
However unlike its other polymorphs of calcite and aragonite, 
when synthesised vaterite will typically form small spherical 
particles making them ideal for optical trapping and rotation. 
Synthesis of Vaterite micro spheres requires fine control of 
the crystal growth process in order to maintain polymorphic 
stability. Though for the purposes of optical rotation the 
exact polymorph is not as important as the morphology as all 
3 polymorphs are inherently birefringent.

Vaterite samples where made by the first preparing equal amounts 
of $CaCl_2$ and $Na_2CO_3$ at a concentration of $0.33M$, at 
the same time a vial of $0.33M\ MgSO_4$ was prepared and set 
aside for later. First a small vial was filled with $1.5mL$ of 
$CaCl_2$ followed by $60\mu L$ and $90\mu L$ of $MgSO_4$ and 
$NaCO_3$ respectively, forming a seed solution. Next, a larger 
vial was filled with $5\ mL$, $1.5\ mL$, and $1\ mL$ of $CaCL_2$, 
$MgSO_4$, and $NaCO_3$ respectively followed by the seed solution. 
After 10 minutes of slow but continuous mixing a few drops of 
Agepon was added to halt the reaction, the solution was filtered 
and washed 3 times with distilled water before being suspended 
in water.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Rotation of birefringent micro spheres} 
Optical tweezing has often been used for micro-rheology, by computing the
exact forces being exerted on the trapped sphere, one can determine the
local temperature/viscosity of the medium \cite{Millen2014, RodriguezSevilla2018}.

When it comes to optical rotation, a rotating particle will experience a
fluid drag torque that is proportional to its rotation rate, the maximum 
rotation rate is when the fluid drag is equal to the optical torque 
of the laser \cite{RodriguezSevilla2018}. Optical rotors have been used
to measure local temperature changes \cite{Millen2014}, or for understanding 
how fluid shear propagates through a medium \cite{Knoener2005}. Likewise, 
one can use a beam steering arrangement to probe the drag force of the 
fluid, by understanding the trap strength (calibrating using a low 
frequency signal) one can measure the drag force experienced by the local 
fluid \cite{RobertsonAnderson2018}. I
